---
titie: "P8130 Homework 4"
author: "Ruiiian Zhang"
date: "11/9/2021"
output: pdf_document
---


```{r, message = FALSE}
library(tidyverse)
```


## Problem 1

Proof: $\sum_{i=1}^{k} \sum_{j=1}^{n_i} (y_{ij}-\bar{\bar{y}})^2 = \sum_{i=1}^{k} \sum_{j=1}^{n_i} (y_{ij}-\bar{y_i})^2 + \sum_{i=1}^{k} \sum_{j=1}^{n_i} (\bar{y_i}-\bar{\bar{y}})^2$  


$\sum_{i=1}^{k} \sum_{j=1}^{n_i} (y_{ij}-\bar{\bar{y}})^2 \\ = \sum_{i=1}^{k} \sum_{j=1}^{n_i} (y_{ij}-\bar{y_i} + \bar{y_i} - \bar{\bar{y}})^2 \\ =  \sum_{i=1}^{k} \sum_{j=1}^{n_i} (y_{ij}-\bar{y_i})^2 + \sum_{i=1}^{k} \sum_{j=1}^{n_i} (\bar{y_i}-\bar{\bar{y}})^2 + 2\sum_{i=1}^{k} \sum_{j=1}^{n_i} (y_{ij} - \bar{y_i}) (\bar{y_i} - \bar{\bar{y}})$

Since $\sum_{j=1}^{n_i} (y_{ij} - \bar{y_i}) = n_i\bar{y_i} - n_i\bar{y_i} = 0$

We have $2\sum_{i=1}^{k} \sum_{j=1}^{n_i} (y_{ij} - \bar{y_i}) (\bar{y_i} - \bar{\bar{y}}) = 0$

Thus $\sum_{i=1}^{k} \sum_{j=1}^{n_i} (y_{ij}-\bar{\bar{y}})^2 = \sum_{i=1}^{k} \sum_{j=1}^{n_i} (y_{ij}-\bar{y_i})^2 + \sum_{i=1}^{k} \sum_{j=1}^{n_i} (\bar{y_i}-\bar{\bar{y}})^2$



## Problem 2

```{r}
crash_df = read_csv("data/crash.csv")
```

### a)

```{r}
descrip_groups = function(group) {
  
  mean = mean(group, na.rm = TRUE)
  range = max(group, na.rm = TRUE) - min(group, na.rm = TRUE)
  sd = sd(group, na.rm = TRUE)
  
  return(
    tibble(
      mean = mean,
      range = range,
      sd = sd
    )
  )
  
}

descrip_groups(crash_df$pedestrian)

descrip_results = map(crash_df, descrip_groups)


bind_rows(
    descrip_results$pedestrian, 
    descrip_results$bicycle,
    descrip_results$car
  ) %>% 
  mutate(
    crash_type = c("pedesterian", "bicycle", "car")) %>% 
  relocate(crash_type) %>% 
  knitr::kable(digits = 3)
```

* Patients with pedestrian crash has the highest average PTSD score of 37.875, compared to the average score of 32.500 in bicycle group and 23.429 in car group. This indicates that patients with pedestrian crash have a highest PTSD level and patients with car crash have a lowest PTSD level.  
* The range of PTSD score in pedestrian group is 14, which is slightly wider than the bicycle group (11) and the car group (11).  
* Patients with pedestrian crash has the highest standard deviation of PTSD score of 5.436, compared to the standard deviation of 4.062 in bicycle group and 3.867 in car group. This indicates that the PTSD score varies more in pedestrian group and varies less in car group.

### b)

$H_0$: all the group means of PTSD score are equal.  
$H_1$: not all the group means of PTSD score are equal.

```{r}
crash_df = 
  pivot_longer(
    data = crash_df,
    cols = pedestrian:car,
    names_to = "crash_type",
    values_to = "ptsd_score"
  )

crash_anova = aov(ptsd_score ~ factor(crash_type), data = crash_df) 

summary(crash_anova)

f_stats = summary(crash_anova)[[1]][["F value"]][1]
f_crit = qf(0.99, 2, 22)
```

$F_{stats} = 19.533$  
$F_{crit} = 5.719$  
$F_{stats} > F_{crit}$  

With a type I error of 0.01, we have enough evidence to reject $H_0$. We can draw a conclusion that not all the group means of PTSD score are equal.

### c)

```{r}
pairwise.t.test(crash_df$ptsd_score, crash_df$crash_type, p.adj = 'bonferroni')

df = 22
qt(0.9967, df)
```

$\alpha^* = 0.01/3 = 0.0033$
$t_{crit} = t_{22, 0.9967} = 2.9996$

The p-value between car and bicycle group is 0.0014, the p-value between pedestrian and bicycle group is 0.0586, the p-value between pedestrian and car group is 9.1e-06, which are all smaller than the adjusted significance level $\alpha^* = 0.0033$. 
Thus we can draw a conclusion that all three groups pairwise have different means from each other.

### d)

Patients with pedestrian crash have the highest PSTD score with a average of 37.875, patients with bicycle crash have a average PTSD score of 32.500, and patients in bicycle group have a average PTSD score of 23.429. Patients with pedestrian crash also has the highest range and standard deviation compared to the other two groups. Based on a type I error of 0.01, the average of PTSD score across different crash type are significantly different, and we found that the average score of each two groups are pairwise different. 


## Problem 3

### a)

We can use chi-squared test for homogeneity, since the data qualifies the following assumptions:  
* Independent random samples  
* No expected cell counts are 0, and no more than 20% of the cells have an expected count less than 5  

### b)

```{r}
drug_data = 
  matrix(
    c(15, 18, 18, 15, 20, 13),
    nrow = 3,
    ncol = 2, 
    byrow = TRUE,
    dimnames = 
      list(c("desipramine", "lithium", "placebo"), c("relapsed", "not_relapsed")))

chisq.test(drug_data)
```

### c)

$H_0$: the proportions of relapse among drug/placebo are equal; and proportions of non-relapse among drug/placebo are equal.  
$H_1$: not all the proportions are equal.  

```{r}
qchisq(0.95, 2)
```

$p-value = 0.4623$
$\chi_{stats} = 1.543$  
$\chi_{crit} = 5.991$  
$\chi_{stats} < \chi_{crit}$  

Base on 5% level of significance, we do not have enough evidence to reject $H_0$. We can draw a conclusion that the proportions of relapse among drug/placebo are equal; and proportions of non-relapse among drug/placebo are equal.

